---
title: "WUP_04"
author: "Joel Calderin"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSS= read.csv("GSScsv.csv")
```
  
# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  

$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  

$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->


Our response variable is "WORDSUM" which is a numerical test result. Our predictor variable is "DEGREE" which is a categorical. We will be using the Numerical~Cat helper file and chatGPT. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

I used chatGPT to assist me with my box plotting. 

```{r}
ggplot(data = GSS , aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot() +
  labs(x = "DEGREE", y = "WORDSUM") +
  ggtitle("Box Plot of WORDSUM vs. DEGREE")
```

This box plot shows us that the degree with the highest median WORDSUM test score is a graduate degree. The lowest median WORDSUM test score is a LT high school degree. Both junior college and bachelor degrees have the same median. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE, data=GSS)
```

According to the chart, we can see that graduate degrees have the highest scoring average at around 10. The bachelor and junior college degrees both have an average around 9. The lower average scores belongs to LT high school and high school degrees at around 8.



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSS)
summary.aov(model1)
```

This shows the P value is less than 0.05, so we reject the null hypothesis. This indicates that there is a difference in WORDSUM tests in comparison to the level of degree. 



```{r}
plot(TukeyHSD(model1))
print(TukeyHSD(model1))
```

We are given evidence that there is a difference in the test scores for the WORDSUM in relation to the level of degree. The higher educational degree an individual had, the higher test score they were likely to get. We know this by seeing if there is an interval of 0 in the testing/graph. If there is not an interval of 0, then there is no difference in the test scores; if there is an interval of 0, then there is a difference in the test scores.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


Is there a difference in test results for the WORDSUM depending on the level of education? Yes, there is a difference in test scores based on education level. The graphical and numerical data display that on average the higher education degree a person had, the higher test score they were likely to receive. We also know there is a difference based on the P-value, that was under 0.05. This means we rejected the NULL hypothesis and determined there is a difference in the data. In conclusion, the test scores for the WORDSUM differed depending on the level of education.